---
title: "TeamWinProbModel"
output: html_document
date: '2022-04-24'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(caret)
library(xgboost)
library(plotly)
```

# Super Schedule 
```{r}
SuperSked08 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2008_super_sked.csv", header = TRUE)
SuperSked09 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2009_super_sked.csv", header = FALSE)
SuperSked10 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2010_super_sked.csv", header = FALSE)
SuperSked11 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2011_super_sked.csv", header = FALSE)
SuperSked12 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2012_super_sked.csv", header = FALSE)
SuperSked13 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2013_super_sked.csv", header = FALSE)
SuperSked14 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2014_super_sked.csv", header = FALSE)
SuperSked15 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2015_super_sked.csv", header = FALSE)
SuperSked16 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2016_super_sked.csv", header = FALSE)
SuperSked17 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2017_super_sked.csv", header = FALSE)
SuperSked18 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2018_super_sked.csv", header = FALSE)
SuperSked19 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2019_super_sked.csv", header = FALSE)
SuperSked20 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2020_super_sked.csv", header = FALSE)
SuperSked21 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2021_super_sked.csv", header = FALSE)
SuperSked22 <- read.csv("C:\\Users\\Jonathan Yu\\Documents\\GitHub\\DS4002Proj2\\Team Game Stats\\2022_super_sked.csv", header = FALSE)
```

# Fix Column names for SuperSked
```{r}
colnames(SuperSked09) <- colnames(SuperSked08)
colnames(SuperSked10) <- colnames(SuperSked08)
colnames(SuperSked11) <- colnames(SuperSked08)
colnames(SuperSked12) <- colnames(SuperSked08)
colnames(SuperSked13) <- colnames(SuperSked08)
colnames(SuperSked14) <- colnames(SuperSked08)
colnames(SuperSked15) <- colnames(SuperSked08)
colnames(SuperSked16) <- colnames(SuperSked08)
colnames(SuperSked17) <- colnames(SuperSked08)
colnames(SuperSked18) <- colnames(SuperSked08)
colnames(SuperSked19) <- colnames(SuperSked08)
colnames(SuperSked20) <- colnames(SuperSked08)
colnames(SuperSked21) <- colnames(SuperSked08)
colnames(SuperSked22) <- colnames(SuperSked08)

```
# Add Binary Result Column to Supersked Function 
```{r}
addbinary <- function(supersked){
  for(row in 1:nrow(supersked)){
    team1 <- supersked[row, "team1"]
    team2 <- supersked[row, "team2"]
    winner <- supersked[row, "winner"]
    if(team1 == winner){
      supersked[row, "binaryresult"] <- 1
    }
    else{
      supersked[row, "binaryresult"] <- 0
    }
  }
  return(supersked)
}
```

# add binary results column to supersked
```{r}
SuperSked08 <- addbinary(SuperSked08)
SuperSked09 <- addbinary(SuperSked09)
SuperSked10 <- addbinary(SuperSked10)
SuperSked11 <- addbinary(SuperSked11)
SuperSked12 <- addbinary(SuperSked12)
SuperSked13 <- addbinary(SuperSked13)
SuperSked14 <- addbinary(SuperSked14)
SuperSked15 <- addbinary(SuperSked15)
SuperSked16 <- addbinary(SuperSked16)
SuperSked17 <- addbinary(SuperSked17)
SuperSked18 <- addbinary(SuperSked18)
SuperSked19 <- addbinary(SuperSked19)
SuperSked20 <- addbinary(SuperSked20)
SuperSked21 <- addbinary(SuperSked21)
SuperSked22 <- addbinary(SuperSked22)

```

# March Madness Games
```{r}
SuperSked08$muid <- as.numeric(SuperSked08$muid)
SuperSked09$muid <- as.numeric(SuperSked09$muid)
SuperSked10$muid <- as.numeric(SuperSked10$muid)
SuperSked11$muid <- as.numeric(SuperSked11$muid)
SuperSked12$muid <- as.numeric(SuperSked12$muid)
SuperSked13$muid <- as.numeric(SuperSked13$muid)
SuperSked14$muid <- as.numeric(SuperSked14$muid)
SuperSked15$muid <- as.numeric(SuperSked15$muid)
SuperSked16$muid <- as.numeric(SuperSked16$muid)
SuperSked17$muid <- as.numeric(SuperSked17$muid)
SuperSked18$muid <- as.numeric(SuperSked18$muid)
SuperSked19$muid <- as.numeric(SuperSked19$muid)
SuperSked20$muid <- as.numeric(SuperSked20$muid)
SuperSked21$muid <- as.numeric(SuperSked21$muid)
SuperSked22$muid <- as.numeric(SuperSked22$muid)

MarchMadness08 <- subset(SuperSked08, muid >= 284000000)
MarchMadness09 <- subset(SuperSked09, muid >= 294000000)
MarchMadness10 <- subset(SuperSked10, (conf == 3 & venue == 1) & (date != "4/1/10" & date != "3/30/10" & muid != 300782250 & date != "3/17/10"))
MarchMadness11 <- subset(SuperSked11, (conf == 3 & venue == 1) & (date != "3/29/11" & muid != 310740103 & muid != 310900333))
MarchMadness12 <- subset(SuperSked12, (conf == 3 & venue == 1) & (muid != 320732168 & date != "3/27/12" & date != "3/29/12"))
MarchMadness13 <- subset(SuperSked13, (conf == 3 & venue == 1) & (muid != 330780096 & date != "4/2/13" & date != "4/4/13"))
MarchMadness14 <- subset(SuperSked14, (conf == 3 & venue == 1)& (date != "4/1/14" & date != "4/3/14"))
MarchMadness15 <- subset(SuperSked15, (conf == 3 & venue == 1) & (date != "4/2/15" & date != "3/31/15" & team1 != "South Dakota St." & team1 != "Evansville" & team1 != "Canisius" & team1 != "Kent St."))
MarchMadness16 <- subset(SuperSked16, (conf == 3 & venue == 1) & (team1 != "Morehead St." & team1 != "Princeton" & date != "3/31/16" & date != "3/30/16" & date != "3/29/16" & date != "3/28/16"))
MarchMadness17 <- subset(SuperSked17, (conf == 3 & venue == 1) & (date != "3/28/17" & date != "3/30/17"))
MarchMadness18 <- subset(SuperSked18, (conf == 3 & venue == 1) & (date != "3/27/18" & date != "3/29/18"))
MarchMadness19 <- subset(SuperSked19, (conf == 3 & venue == 1) & (date != "4/2/19" & date != "4/4/19"))
MarchMadness21 <- subset(SuperSked21, (conf == 3 & venue == 1) & (date != "3/17/21" & team1  != "Drake" & team1 != "North Carolina St." & team1 != "SMU" & team1 != "Buffalo" & date != "3/25/21" & team1 != "Mississippi St." & team1 != "Louisiana Tech" & team1 != "Memphis" & date != "3/24/21" & date != "3/23/21" & team1 != "Bellarmine" & team1 != "Stetson" & team1 != "Longwood" & team2 != "Bryant"))
MarchMadness22 <- subset(SuperSked22, (conf == 3 & venue == 1) & date != "3/31/22" & date != "3/29/22" & date != "3/21/22" & date != "3/22/22" & date != "3/23/22" & team1 != "Boston University" & team1 != "Florida Atlantic" & team1 != "Abilene Christian" & team1 != "UNC Wilmington" & team1 != "Ohio" & team1 != "Stephen F. Austin" & team1 != "Drake" & team1 != "Cal Baptist")
```
# Save These Columns for Analysis
```{r}
savecolumns <- function(marchmadness){
  marchmadness <- marchmadness[, c("date", "conmatch", "matchup", "team1", "team2", "binaryresult")]
  return(marchmadness)
}
MarchMadness08 <- savecolumns(MarchMadness08)
MarchMadness09 <- savecolumns(MarchMadness09)
MarchMadness10 <- savecolumns(MarchMadness10)
MarchMadness11 <- savecolumns(MarchMadness11)
MarchMadness12 <- savecolumns(MarchMadness12)
MarchMadness13 <- savecolumns(MarchMadness13)
MarchMadness14 <- savecolumns(MarchMadness14)
MarchMadness15 <- savecolumns(MarchMadness15)
MarchMadness16 <- savecolumns(MarchMadness16)
MarchMadness17 <- savecolumns(MarchMadness17)
MarchMadness18 <- savecolumns(MarchMadness18)
MarchMadness19 <- savecolumns(MarchMadness19)
MarchMadness21 <- savecolumns(MarchMadness21)
MarchMadness22 <- savecolumns(MarchMadness22)
```

# March Madness modelData by Lee
```{r}
AllData <- read.csv("C:\\Users\\Jonathan Yu\\Downloads\\modelData.csv")
```

# March Madness Merging 
```{r}
MarchMadness09 <- merge(AllData, MarchMadness09, by.x = "matchup", by.y = "matchup")
MarchMadness10 <- merge(AllData, MarchMadness10, by.x = "matchup", by.y = "matchup")
MarchMadness11 <- merge(AllData, MarchMadness11, by.x = "matchup", by.y = "matchup")
MarchMadness12 <- merge(AllData, MarchMadness12, by.x = "matchup", by.y = "matchup")
MarchMadness13 <- merge(AllData, MarchMadness13, by.x = "matchup", by.y = "matchup")
MarchMadness14 <- merge(AllData, MarchMadness14, by.x = "matchup", by.y = "matchup")
MarchMadness15 <- merge(AllData, MarchMadness15, by.x = "matchup", by.y = "matchup")
MarchMadness16 <- merge(AllData, MarchMadness16, by.x = "matchup", by.y = "matchup")
MarchMadness17 <- merge(AllData, MarchMadness17, by.x = "matchup", by.y = "matchup")
MarchMadness18 <- merge(AllData, MarchMadness18, by.x = "matchup", by.y = "matchup")
MarchMadness19 <- merge(AllData, MarchMadness19, by.x = "matchup", by.y = "matchup")
MarchMadness21 <- merge(AllData, MarchMadness21, by.x = "matchup", by.y = "matchup")
```

```{r}
march_training <- bind_rows(MarchMadness09, MarchMadness10, MarchMadness11,MarchMadness12,MarchMadness13,MarchMadness14,MarchMadness15,MarchMadness16,MarchMadness17,MarchMadness18,MarchMadness19,MarchMadness21)
march_training <- march_training[,51:220]
march_training <- march_training[, -c(1, 4, 83, 86, 96, 14)]
march_training$T1conf <- as.numeric(as.factor(march_training$T1conf))
march_training$T2conf <- as.numeric(as.factor(march_training$T2conf))
training_labels <- as.matrix(march_training$binaryresult)
matrix_march_training <- as.matrix(march_training[,-c(159:164)])
```

# XGBoost Model 1
```{r}
set.seed(132)
march_model <- xgboost(data = matrix_march_training, label = as.matrix(training_labels), max.depth = 2, eta = 0.3, nthread = 2, nrounds = 500, model = "binary:logistic")

```

# Test on Testing Data
```{r}
AllData2022 <- read.csv("C:\\Users\\Jonathan Yu\\Downloads\\modelData2022.csv")
MarchMadness22 <- merge(AllData2022, MarchMadness22, by.x = "matchup", by.y = "matchup")
march_trainingcols <- march_training[,-c(159:164)]
march_testing <- MarchMadness22
march_testing <- select(march_testing, c(colnames(march_trainingcols)))
march_testing$T1conf <- as.numeric(as.factor(march_testing$T1conf))
march_testing$T2conf <- as.numeric(as.factor(march_testing$T2conf))
pred <- predict(march_model, as.matrix(march_testing))
predf <- as.data.frame(pred)
predf$real <- MarchMadness22$binaryresult
predf$error <- abs(predf$real - predf$pred)
modelanalysis <- data.frame(row = 0)
modelanalysis$mse <- mean((predf$real-predf$pred)^2)
modelanalysis$mae <- MAE(predf$real, predf$pred)
modelanalysis$RMSE <- caret::RMSE(predf$real, predf$pred)
pred <- as.numeric(pred>0.5)
pred
confusionMatrix(factor(pred), factor(MarchMadness22$binaryresult))
modelanalysis
save(march_model, file = "march_model_1.RData")
```
# Importance Matrix
```{r}
march_testing
importance_matrix <- xgb.importance(model = march_model)
xgb.plot.importance(importance_matrix, top_n = 5)
```
# Second Model Attempt with less features...
```{r}
march_training2 <- march_training[,c("T1adjoe", "T1conf", "T1adjde", "T1sos", "T1ncsos", "T1consos", "T1Opp.OE", "T1Opp.DE", "T1Con.Adj.OE", "T1Con.Adj.DE","T1winRate", "T1eFG.", "T1eFG..Def", "T1FTR", "T1FTR.Def", "T1TO.", "T1TO..Def.", "T13P.", "T13pD.", "T1ft.", "T12p.D", "T1ft.D", "T13P.rate", "T13P.rate.D", "T1arate", "T1arateD", "T2adjoe", "T2conf", "T2adjde", "T2sos", "T2ncsos", "T2consos", "T2Opp.OE", "T2Opp.DE", "T2Con.Adj.OE", "T2Con.Adj.DE","T2winRate", "T2eFG.", "T2eFG..Def", "T2FTR", "T2FTR.Def", "T2TO.", "T2TO..Def.", "T23P.", "T23pD.", "T2ft.", "T22p.D", "T2ft.D", "T23P.rate", "T23P.rate.D", "T2arate", "T2arateD")]

march_model_2 <- xgboost(data = as.matrix(march_training2), label = as.matrix(training_labels), max.depth = 2, eta = 0.3, nthread = 2, nrounds = 500, model = "binary:logistic")

march_testing2 <- march_testing[,c("T1adjoe", "T1conf", "T1adjde", "T1sos", "T1ncsos", "T1consos", "T1Opp.OE", "T1Opp.DE", "T1Con.Adj.OE", "T1Con.Adj.DE","T1winRate", "T1eFG.", "T1eFG..Def", "T1FTR", "T1FTR.Def", "T1TO.", "T1TO..Def.", "T13P.", "T13pD.", "T1ft.", "T12p.D", "T1ft.D", "T13P.rate", "T13P.rate.D", "T1arate", "T1arateD", "T2adjoe", "T2conf", "T2adjde", "T2sos", "T2ncsos", "T2consos", "T2Opp.OE", "T2Opp.DE", "T2Con.Adj.OE", "T2Con.Adj.DE","T2winRate", "T2eFG.", "T2eFG..Def", "T2FTR", "T2FTR.Def", "T2TO.", "T2TO..Def.", "T23P.", "T23pD.", "T2ft.", "T22p.D", "T2ft.D", "T23P.rate", "T23P.rate.D", "T2arate", "T2arateD")]

pred2 <- predict((march_model_2), as.matrix(march_testing2))
pred2 <- as.numeric(pred2 > 0.5)
confusionMatrix(factor(pred2), factor(MarchMadness22$binaryresult))
importance_matrix2 <- xgb.importance(model = march_model_2)
xgb.plot.importance(importance_matrix2, top_n = 5)
save(march_model_2, file = "march_model_2.RData")
```
# Third model with even less features
```{r}
march_training3 <- march_training2[,c("T1adjoe", "T1adjde", "T1sos", "T1winRate", "T2adjoe", "T2adjde", "T2sos", "T2winRate")]

march_model_3 <- xgboost(data = as.matrix(march_training3), label = as.matrix(training_labels), max.depth = 2, eta = 0.3, nthread = 2, nrounds = 500, model = "binary:logistic")
march_testing3 <- march_testing2[,c("T1adjoe", "T1adjde", "T1sos", "T1winRate", "T2adjoe", "T2adjde", "T2sos", "T2winRate")]
pred3 <- predict(march_model_3, as.matrix(march_testing3))
pred3 <- as.numeric(pred3 > 0.5)
confusionMatrix(factor(pred3), factor(MarchMadness22$binaryresult))
importance_matrix3 <- xgb.importance(model = march_model_3)
xgb.plot.importance(importance_matrix3, top_n = 8)
save(march_model_3, file = "march_model_3.RData")
```
```{r}
readRDS("diff_xg_model.RData")
```


